---
title: "Recorded Fish Passage at Willamette Falls, OR"
author: "Carlos Simms"
date: "2/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE) # To suppress code messages and warnings.
library(tidyverse)
library(lubridate)
library(here)
library(tsibble)
library(feasts)
library(slider)
library(gghighlight)
library(patchwork)
library(janitor) # The only added package.
```

```{r}
salmon <- read.csv(here("data", 
                        "willamette_fish_passage.csv")) %>% 
  clean_names() # To clean the data set.

salmon_ts <- salmon %>% 
  mutate(date = mdy(date)) %>% 
  as_tsibble(key = NULL, 
             index = date) %>% 
  pivot_longer(cols = 3:15, 
               names_to = "species", 
               values_to = "count") %>% 
  filter(species %in% c("coho", 
                        "jack_coho", 
                        "steelhead")) %>% 
  replace_na(list(count = 0)) # To change the original data code to a cleaned, vertical structure.
```

```{r}
annual_totals <- salmon_ts %>%
  mutate(year = year(date), 
         .keep = "unused") %>% # To mutate the lubridate::date as a new variable "year" and omit the source variable "date".
  select(species, 
         count, 
         year) %>% # To select the variables of interest.
  group_by(year, 
           species) %>% 
  summarize(annual_total = sum(count)) # To group by year and species to derive annual totals.
```

```{r}
facet_names <- c('coho' = "Coho", 
                 'jack_coho' = "Jack Coho", 
                 'steelhead' = "Steelhead") # To change display names of the species via a vector.

ggplot(data = annual_totals, 
       aes(x = year, 
           y = annual_total)) + 
  geom_col(aes(fill = species), 
           show.legend = FALSE) + 
  facet_wrap(~species) + # OPINION: The best way to visualize this was via a faceted column graph.
  labs(title = "Total Annual Passage of Salmon Species From 2001 to 2010",
       x = "Year", 
       y = "Total Annual Fish Passage") + 
  theme_bw() + 
  theme(plot.title = element_text(face = "bold", 
                                  hjust = 0.5), 
        axis.title.x = element_text(vjust = -2), 
        axis.title.y = element_text(vjust = 3)) + 
  facet_grid(. ~ species, 
             labeller = as_labeller(facet_names)) + # To specify names as the vector.
  scale_x_continuous(breaks = c(2001:2010), 
                     labels = c("01", "02", "03", "04", "05", 
                                "06", "07", "08", "09", "10")) # To rename X axis value labels to simpler indeces.
```

**Figure 3.** Total observed annual salmon passage count by species, between 2001 and 2010 via fish ladders at Willamette Falls, Oregon. 

- More Steelhead Salmon have passed through the Willamette Falls fish ladders than that of any of the other two species.

- There was a significant increased count of Coho Salmon passing through the Willamette Falls fish ladders in 2009 and 2010, versus in any of the previous record years.

- There is a decreasing trend in Steelhead Salmon count and a consistent trend of Jack Coho Salmon count passing through the Willamette Falls fish ladders throughout the 2001 to 2010 period.